Stochastic Variance Reduced Policy Gradient

### 2018-01-18 svrpg.tex
done writing first proposition of Lemma 1

\section{Definitions}
We give some additional definitions which will be useful in the proofs of Appendix \ref{app:proofs}.
\begin{definition}
For a random variable $X$:
\begin{align*}
	&\mathbb{E}_{t\mid s}\left[X\right] \coloneqq 
		\mathop{\mathbb{E}}_{\substack{\tau_j\sim\pi(\cdot\mid\tilde{\vtheta}^s) \\ \tau_i\sim\pi(\cdot\mid\vtheta^{s+1}_i)\text{ for $i=0,\dots,t-1$}}}{\left[X \mid \tilde{\vtheta^s}\right]} \\
	&\coloneqq \EVV[\tau_j\sim\pi(\cdot\mid\tilde{\vtheta}^s)]{
			\EVV[\tau_0\sim\pi(\cdot\mid\vtheta_0^{s+1})]
				{\dots
					\EVV[\tau_t\sim\pi(\cdot\mid\vtheta_t^{s+1})]
						{X\mid\vtheta_t^{s+1}}
				 \dots
			\mid\vtheta_0^{s+1}}
		\mid\tilde{\vtheta}^s},
\end{align*}
where the sequence $\tilde{\vtheta}^s,\vtheta_0^{s+1},\dots,\vtheta_t^{s+1}$ is defined in Algorithm \ref{alg:svrpg}. 
\end{definition}
Intuitively, the $\Ets{\cdot}$ operator computes the expected value with respect to the sampling of trajectories from the snapshot $\tilde{\vtheta}^s$ up to the $t$-th iteration included. Note that the order in which expected values are taken is important since each $\vtheta_{t}^{s+1}$ is function of previously sampled trajectories and is used to sample new ones.

@@

@@

\convergence*
\begin{proof}
We prove the theorem for the following values of the constants:
\begin{align*}
& \psi \coloneqq \min_t\Psi_t, \\
& D \coloneqq \max_t d_t \\
& c_m \coloneqq 0 \\
& c_t \coloneqq c_{t+1}(1+\alpha_t\beta+3\alpha_t^2L^2), \\
\end{align*}
where:
\begin{align*}
	& \beta > 0 \\
	& \Psi_t \coloneqq (\alpha_t - \frac{c_{t+1}\alpha_t}{\beta} - \frac{3}{2}\alpha_t^2L - 3c_{t+1}\alpha_t^2) \\
	& d_t \coloneqq 3(\alpha_t\frac{L}{2} + c_{t+1}\alpha_t^2), \\
\end{align*}
and $\beta,\alpha_t$ are chosen so that $\psi_t > 0$ for $t=1,\dots,m$.
We use two auxiliary lemmas, whose proofs are provided later on:
%Lemma 1
\begin{restatable}[]{lemma}{auxone}\label{lemma:aux1}
\[
	\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
	\frac{R_{t+1}^{s+1} - R_t^{s+1} + Z_t}{\Psi_t},
\]
where
\[
	R_t^{s+1}\coloneqq \Ets{J(\vtheta_t^{s+1}) + c_t\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
\]
\end{restatable}
%Lemma 2
\begin{restatable}[]{lemma}{auxtwo}\label{lemma:aux2}
	...
\end{restatable}

\end{proof}

%Proof of Lemma 1
\auxone*
\begin{proof}
We have:
\begin{align}
	\Ets{J(\vtheta_{t+1}^{s+1})} 
	&\geq \Ets{J(\vtheta_t^{s+1})+\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}} - \frac{L}{2}\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}}^2} \\
	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradBlack{\vtheta_t^{s+1}}} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
	&\geq \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradIdeal{\vtheta_t^{s+1}}+e_s} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2}  + \alpha_t\Ets[t-1]{\dotprod{e_s}{\gradJ{\vtheta_t^{s+1}}}} \\
	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} +
	\alpha_t\Covts[t-1]{e_s}{\gradJ{\vtheta_t^{s+1}}} \\
	&\geq  \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} -
	\alpha_t\left|\Covts[t-1]{e_s}{\gradJ{\vtheta_t^{s+1}}}\right| \\
	&\geq  \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
	&+
	\alpha_t\frac{\sqrt{\Varts[]{\gradApp{\tilde{\vtheta}^s}{1}}\Varts{\gradJ{\vtheta_t^{s+1}}}}}{\sqrt{N}}
\end{align}
where (1) is from the L-smoothness of $J(\vtheta)$, (2) is from the SVRPG gradient ascent update, (3) is from Definition \ref{def:ideal}, (4) is from the fact that $\EVV[\tau_t\sim\pi(\cdot\mid \vtheta_t^{s+1})]{\gradBlack{\vtheta_t^{s+1}}\mid\vtheta_t^{s+1}} = \EVV[\tau_t\sim\pi(\cdot\mid \vtheta_t^{s+1})]{\gradIdeal{\vtheta_t^{s+1}}\mid\vtheta_t^{s+1}}$, (5) is from the fact that $\Ets[t-1]{e_s} = 0$, and (7)+(8) is from the definition of $\gradApp{\vtheta}{N}$.

@@

>>>>

\subsection*{Definitions}
We give some additional definitions which will be useful in the proofs.
\begin{definition}
For a random variable $X$:
\begin{align*}
	&\mathbb{E}_{t\mid s}\left[X\right] \coloneqq 
		\mathop{\mathbb{E}}_{\substack{\tau_j\sim\pi(\cdot\mid\tilde{\vtheta}^s) \\ \tau_i\sim\pi(\cdot\mid\vtheta^{s+1}_i)\text{ for $i=0,\dots,t-1$}}}{\left[X \mid \tilde{\vtheta^s}\right]} \\
	&\coloneqq \EVV[\tau_j\sim\pi(\cdot\mid\tilde{\vtheta}^s)]{
			\EVV[\tau_0\sim\pi(\cdot\mid\vtheta_0^{s+1})]
				{\dots
					\EVV[\tau_t\sim\pi(\cdot\mid\vtheta_t^{s+1})]
						{X\mid\vtheta_t^{s+1}}
				 \dots
			\mid\vtheta_0^{s+1}}
		\mid\tilde{\vtheta}^s},
\end{align*}
where the sequence $\tilde{\vtheta}^s,\vtheta_0^{s+1},\dots,\vtheta_t^{s+1}$ is defined in Algorithm \ref{alg:svrpg}. 
\end{definition}
Intuitively, the $\Ets{\cdot}$ operator computes the expected value with respect to the sampling of trajectories from the snapshot $\tilde{\vtheta}^s$ up to the $t$-th iteration included. Note that the order in which expected values are taken is important since each $\vtheta_{t}^{s+1}$ is function of previously sampled trajectories and is used to sample new ones.

@@

\subsection*{Lemmas}
Before addressing the main convergence theorem, we prove some useful lemmas.

@@

%Lemma 1
\begin{restatable}[]{lemma}{auxone}\label{lemma:aux1}
	\[
	\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
	\frac{R_{t+1}^{s+1} - R_t^{s+1} + Z_t}{\Psi_t},
	\]
	where
	\[
	R_t^{s+1}\coloneqq \Ets{J(\vtheta_t^{s+1}) + c_t\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}
	\]
\end{restatable}
\begin{proof}
	We have:
	\begin{align}
	\Ets{J(\vtheta_{t+1}^{s+1})} 
	&\geq \Ets{J(\vtheta_t^{s+1})+\dotprod{\gradJ{\vtheta_t^{s+1}}}{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}} - \frac{L}{2}\norm[]{\vtheta_{t+1}^{s+1}-\vtheta_t^{s+1}}^2} \\
	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradBlack{\vtheta_t^{s+1}}} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
	&\geq \Ets{J(\vtheta_t^{s+1})+\alpha_t\dotprod{\gradJ{\vtheta_t^{s+1}}}{\gradIdeal{\vtheta_t^{s+1}}+e_s} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2}  + \alpha_t\Ets[t-1]{\dotprod{e_s}{\gradJ{\vtheta_t^{s+1}}}} \nonumber\\
	&= \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \\
	&\qquad+
	\alpha_t\Tr\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{\gradJ{\vtheta_t^{s+1}}} \\
	&\geq  \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
	&\qquad-
	\alpha_t\Tr\left|\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{\gradJ{\vtheta_t^{s+1}}}\right| \nonumber \\
	%
	&\geq  \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
	&\qquad-\alpha_t \Tr\left(\sqrt{\Varts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{\gradJ{\vtheta_t^{s+1}}}}\right) \nonumber \\
	%
	&\geq  \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
	&\qquad-\frac{\alpha_t}{N} \Tr\left(\sqrt{\Varts[t-1]{\gradApp{\tilde{\vtheta}^s}{1}}}\circ\sqrt{\Varts[t-1]{\gradJ{\vtheta_t^{s+1}}}}\right) \\
	%
	&=
	\Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
	&\qquad-\frac{\alpha_t}{2N}\left(\Tr\Varts[t-1]{\gradApp{\tilde{\vtheta}^s}{1}} + \Tr\Varts[t-1]{\gradJ{\vtheta_t^{s+1}}}\right) \\
	%
	&\geq  \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
	&\qquad-\frac{\alpha_t}{2N}\left(\Tr\Varts[t-1]{\gradApp{\tilde{\vtheta}^s}{1}} + \Ets[t-1]{\norm[]{\gradJ{\vtheta_t^{s+1} - \gradJ{\tilde{\vtheta}^s}}}^2}\right) \nonumber\\
	%
	&\geq  \Ets{J(\vtheta_t^{s+1})+\alpha_t\norm[]{\gradJ{\vtheta_t^{s+1}}^2} - \frac{\alpha_t^2L}{2}\norm[]{\gradBlack{\vtheta_t^{s+1}}}^2} \nonumber\\
	&\qquad-\frac{\alpha_tL^2}{2N}\left(\Tr\Varts[t-1]{\gradApp{\tilde{\vtheta}^s}{1}} + \Ets[t-1]{\norm[]{\vtheta_t^{s+1} - \tilde{\vtheta}^s}^2}\right)
\end{align}
where (1) is from the L-smoothness of $J(\vtheta)$, (2) is from the SVRPG gradient ascent update, (3) is from Definition \ref{def:ideal}, (4) is from the fact that $\EVV[\tau_t\sim\pi(\cdot\mid \vtheta_t^{s+1})]{\gradBlack{\vtheta_t^{s+1}}\mid\vtheta_t^{s+1}} = \EVV[\tau_t\sim\pi(\cdot\mid \vtheta_t^{s+1})]{\gradIdeal{\vtheta_t^{s+1}}\mid\vtheta_t^{s+1}}$, (5) is from the fact that $\Ets[t-1]{e_s} = 0$, (6) is from the definition of $\gradApp{\vtheta}{N}$, (7) is from Young's inequality and (8) is again from L-smoothness.

@@

%Lemma 2
\begin{restatable}[]{lemma}{auxtwo}\label{lemma:aux2}
	...
\end{restatable}

\subsection*{Main theorem}

\convergence*
\begin{proof}
We prove the theorem for the following values of the constants:
\begin{align*}
& \psi \coloneqq \min_t\Psi_t, \\
& D \coloneqq \max_t d_t \\
& c_m \coloneqq 0 \\
& c_t \coloneqq c_{t+1}(1+\alpha_t\beta+3\alpha_t^2L^2), \\
\end{align*}
where:
\begin{align*}
	& \beta > 0 \\
	& \Psi_t \coloneqq (\alpha_t - \frac{c_{t+1}\alpha_t}{\beta} - \frac{3}{2}\alpha_t^2L - 3c_{t+1}\alpha_t^2) \\
	& d_t \coloneqq 3(\alpha_t\frac{L}{2} + c_{t+1}\alpha_t^2), \\
\end{align*}
and $\beta,\alpha_t$ are chosen so that $\psi_t > 0$ for $t=1,\dots,m$.
\end{proof}
